Predicting Game of the Year (GOTY) 2024: A Bayesian Analysis of Video Game Metrics from Metacritic

### 2024-11-26 paper/paper.qmd
update paper

---
title: "Predicting The Game Award for Game of the Year 2024: Analyzing Trends and Ratings in the Video Game Industry (2014–2024)"
subtitle: "A Logistic Regression Approach Using User and Critic Data on Metacritic to Forecast the 2024 GOTY Winner"
author: 
  - Ziqi Zhu
thanks: "Code and data are available at: [https://github.com/zzq20010617/2024TGA-goty-predictions](https://github.com/zzq20010617/2024TGA-goty-predictions)."
date: today
date-format: long
abstract: ""
number-sections: true
bibliography: references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

@@

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`, reflecting no strong assumptions about the weight of each predictors' effects.

@@

To get a prediction by the Bayesian model, we use games that are released in 2024 and generate posterior predictions by the new data frame. The mean probability (mean_prob) was used as the primary metric for comparison because the data contains significant class imbalance (10 winner out of 3993 games for past decade). Given that only a small proportion of games achieve GOTY winner, the model is more likely to assign low probabilities for the majority of games, resulting in most predictions being 0 if using a hard classification threshold like 0.5. By using the mean_prob, it allows us to rank games based on their likelihood of being the winners rather than forcing a binary classification, which are more informative for identifying potential GOTY contenders, especially when making comparisons within the 2024 dataset. In addition, I selected the top 10th percentile of games based on their mean_prob and set their goty_status to 1, marking them as potential winners of GOTY. The results are displayed in @tbl-winner. For example, the first row shows that "Elden Ring: Shadow of the Erdtree" has a mean_prob of 0.06675, indicating that it is predicted to win the GOTY award in 267 out of 4000 iterations.

>>>>

---
title: "Predicting Game of the Year (GOTY) 2024: A Bayesian Analysis of Video Game Metrics from Metacritic"
subtitle: "A Logistic Regression Approach Analysing Data from 2014–2023 to Forecast GOTY Success at The Game Awards"
author: 
  - Ziqi Zhu
thanks: "Code and data are available at: [https://github.com/zzq20010617/2024TGA-goty-predictions](https://github.com/zzq20010617/2024TGA-goty-predictions)."
date: today
date-format: long
abstract: "Abstract..."
number-sections: true
bibliography: references.bib
format:
  pdf:
    toc: true
    number-sections: true
    colorlinks: true
---

@@

I run the model in R [@citeR] using the `rstanarm` package of @rstanarm. I use the default priors from `rstanarm`, reflecting no strong assumptions about the weight of each predictors' effects.

@@

To get a prediction by the Bayesian model, I use games that are released in 2024 and generate posterior predictions by the new data frame. The mean probability (mean_prob) was used as the primary metric for comparison because the data contains significant class imbalance (10 winner out of 3993 games for past decade). Given that only a small proportion of games achieve GOTY winner, the model is more likely to assign low probabilities for the majority of games, resulting in most predictions being 0 if using a hard classification threshold like 0.5. By using the mean_prob, it allows us to rank games based on their likelihood of being the winners rather than forcing a binary classification, which are more informative for identifying potential GOTY contenders, especially when making comparisons within the 2024 dataset. In addition, I selected the top 10th percentile of games based on their mean_prob and set their goty_status to 1, marking them as potential winners of GOTY. The results are displayed in @tbl-winner. For example, the first row shows that "Elden Ring: Shadow of the Erdtree" has a mean_prob of 0.06675, indicating that it is predicted to win the GOTY award in 267 out of 4000 iterations.
